\section*{1.4 Mission Requirements}
\addcontentsline{toc}{section}{1.4 Mission Requirements}

The mission requirements are derived directly from the science and exploration objectives, the customer-imposed constraints in Table 1 of the Mission Task Document, and the anticipated operational environment on the Martian surface. These requirements provide measurable criteria by which the success of the mission can be assessed and guide system design and validation.

\textbf{Science Requirements:}
\begin{itemize}
    \item The rover shall detect and characterize subsurface water ice and hydrated minerals within the top 1 meter of regolith.
    \item The rover shall collect data at a minimum of two locations within the 10 km $\times$ 10 km Exploration Zone (EZ).
    \item The rover shall provide in situ analysis of physical and chemical properties at each measurement site.
    \item The rover shall include an astrobiology payload with a volume not exceeding 0.5 m$^3$ and a mass not exceeding 15 kg.
\end{itemize}

\textbf{Exploration Requirements:}
\begin{itemize}
    \item The rover shall assess the accessibility and state of preservation of subsurface water ice for ISRU purposes.
    \item The rover shall identify locations that meet NASA criteria for future crewed mission support.
\end{itemize}

\textbf{Engineering Constraints:}
\begin{itemize}
    \item Total system mass shall not exceed 200 kg at launch.
    \item The stowed volume shall fit within a cube of 2.5 meters per side.
    \item The power system shall not include radioisotope sources and shall rely on solar or alternative non-nuclear energy.
    \item The total mission cost shall not exceed \$450 million.
\end{itemize}

\textbf{Operational Requirements:}
\begin{itemize}
    \item The rover shall traverse a minimum of 5 km during the mission to ensure access to multiple sites.
    \item The rover shall survive and operate in typical mid-latitude Martian surface conditions, including expected temperature extremes, dust exposure, and terrain variability.
    \item The system shall store and transmit science data with sufficient resolution and fidelity to support mission goals.
\end{itemize}

These requirements ensure that the science objectives can be achieved while maintaining compliance with mass, volume, cost, and environmental constraints specified by NASA. They also support strategic planning for future human exploration of Mars by generating actionable data on water ice accessibility and site suitability.

